HO unification from object language to meta language

### 2024-03-20 paper/appendix.tex
appendix add progress

>>>>

\begin{elpicode}

  macro @one :- s z.

  type contract-rigid list ho.tm -> ho.tm -> ho.tm -> o.
  contract-rigid L (ho.lam F) T :- 
    pi x\ contract-rigid [x|L] (F x) T. % also checks H Prefix does not see x
  contract-rigid L (ho.app [H|Args]) T :- 
    std.rev L LRev, std.appendR Prefix LRev Args,
    if (Prefix = []) (T = H) (T = ho.app [H|Prefix]).

  type progress-eta-link ho.tm -> ho.tm -> ho.subst -> ho.subst -> links -> o.
  progress-eta-link (ho.app _ as T) (ho.lam x\ _ as T1) H H1 [] :- !, 
    ({eta-expand T @one} ==l T1) H H1.
  progress-eta-link (ho.con _ as T) (ho.lam x\ _ as T1) H H1 [] :- !, 
    ({eta-expand T @one} ==l T1) H H1.
  progress-eta-link (ho.lam _ as T) T1 H H1 [] :- !, 
    (T ==l T1) H H1.
  progress-eta-link (ho.uva _ _ as X) T H H1 [] :- 
    contract-rigid [] T T1, !, (X ==l T1) H H1.
  progress-eta-link (ho.uva Ad _ as T1) T2 H H [@val-link-eta T1 T2] :- !, 
    if (ho.not_occ Ad H T2) true fail.

  type is-in-pf ho.tm -> o.
  is-in-pf (ho.app [ho.uva _ _ | _]) :- !, fail.
  is-in-pf (ho.lam B) :- !, pi x\ is-in-pf (B x).
  is-in-pf (ho.con _).
  is-in-pf (ho.app L) :- forall1 is-in-pf L.
  is-in-pf N :- name N.
  is-in-pf (ho.uva _ L) :- pattern-fragment L.

  type arity ho.tm -> nat -> o.
  arity (ho.con _) z.
  arity (ho.app L) A :- len L A.

  type occur-check-err ho.tm -> ho.tm -> ho.subst -> o.
  occur-check-err (ho.con _) _ _ :- !.
  occur-check-err (ho.app _) _ _ :- !.
  occur-check-err (ho.lam _) _ _ :- !.
  occur-check-err (ho.uva Ad _) T S :-
    not (ho.not_occ Ad S T).

  type progress-beta-link-aux ho.tm -> ho.tm -> 
          ho.subst -> ho.subst -> links -> o.
  progress-beta-link-aux T1 T2 S1 S2 [] :-  is-in-pf T2, !,
    (T1 ==l T2) S1 S2.
  progress-beta-link-aux T1 T2 S S [@val-link-beta T1 T2] :- !.

  type progress-beta-link ho.tm -> ho.tm -> ho.subst -> 
        ho.subst -> links -> o.
  progress-beta-link T (ho.app[ho.uva V Scope | L] as T2) S S2 [@val-link-beta T T2] :- 
    arity T Arity, len L ArgsNb, ArgsNb >n Arity, !,
    minus ArgsNb Arity Diff, mem.new S V1 S1,
    eta-expand (ho.uva V1 Scope) Diff T1,
    ((ho.uva V Scope) ==l T1) S1 S2.

  progress-beta-link (ho.uva _ _ as T) (ho.app[ho.uva Ad1 Scope1 | L1] as T1) S1 S3 NewLinks :-
    std.append Scope1 L1 Scope1L,
    split-pf Scope1L Scope2 L2,
    not (Scope1 = Scope2), !,
    mem.new S1 Ad2 S2,
    len Scope1 Scope1Len,
    len Scope2 Scope2Len,
    make-eta-link Scope1Len Scope2Len Ad1 Ad2 [] LinkEta S2 S3,
    if (L2 = []) (NewLinks = LinkEta, T2 = ho.uva Ad2 Scope2) 
      (T2 = ho.app [ho.uva Ad2 Scope2 | L2], 
      NewLinks = [@val-link-beta T T2 | LinkEta]).

  progress-beta-link T1 (ho.app[ho.uva _ _ | _] as T2) _ _ _ :- 
    not (T1 = ho.uva _ _), !, fail.

  progress-beta-link (ho.uva _ _ as T) (ho.app[ho.uva _ _ | _] as T2) S1 _ _ :- 
    occur-check-err T T2 S1, !, fail.

  progress-beta-link T1 (ho.app[ho.uva _ _ | _] as T2) H H [@val-link-beta T1 T2] :- !.

  progress-beta-link T1 (ho.app [Hd | Tl]) S1 S2 B :-
    ho.beta Hd Tl T3, 
    progress-beta-link-aux T1 T3 S1 S2 B.

  type solve-link-abs link -> links -> ho.subst -> ho.subst -> o.
  solve-link-abs (ho.abs X) R H H1 :- 
    pi x\ ho.copy x x => (pi S\ ho.deref S x x) => 
      solve-link-abs (X x) (R' x) H H1,
    close-links R' R.

  solve-link-abs (@val-link-eta A B) NewLinks S S1 :- !,
    progress-eta-link A B S S1 NewLinks.

  solve-link-abs (@val-link-beta A B) NewLinks S S1 :- !,
    progress-beta-link A B S S1 NewLinks.

  type take-link link -> links -> link -> links -> o.
  take-link A [B|XS] B XS :- link-abs-same-lhs A B, !.
  take-link A [L|XS] B [L|YS] :- take-link A XS B YS.

  type link-abs-same-lhs link -> link -> o.
  link-abs-same-lhs (ho.abs F) B :- 
    pi x\ link-abs-same-lhs (F x) B.
  link-abs-same-lhs A (ho.abs G) :- 
    pi x\ link-abs-same-lhs A (G x).
  link-abs-same-lhs (@val-link-eta (ho.uva N _) _) (@val-link-eta (ho.uva N _) _).

  type solve-links links -> links -> ho.subst -> ho.subst -> o.
  solve-links [] [] X X.
  solve-links [A|L1] [A|L3] S S2 :- take-link A L1 B L2, !,
    same-link-eta A B S S1, 
    solve-links L2 L3 S1 S2.
  solve-links [L0|L1] L3 S S2 :- deref-link S L0 L,
    solve-link-abs L R S S1, !,
    solve-links L1 L2 S1 S2, std.append R L2 L3.
\end{elpicode}
